# LaTeX

## What is LaTeX?  {.flexbox .vcenter}

- Pronounced «Lah-tech» or «Lay-tech» (to rhyme with «Bertolt Brecht»)
- A document preparation system for high-quality typesetting 
- It is most often used for medium-to-large technical or scientific documents
- Can be used for almost any form of publishing.
- Typesetting journal articles, technical reports, books, and slide presentations
- Allows for precise mathematical statements
- https://www.latex-project.org

## What is LaTeX?  {}

- LaTeX is not a word processor! 
- LaTeX encourages authors not to worry too much about the appearance of their documents but to concentrate on getting the right content. 
- Control over large documents containing sectioning, cross-references, tables and figures.
- Typesetting of complex mathematical formulas.
- Automatic generation of bibliographies and indexes.
- Multi-lingual typesetting
- https://bookdown.org/yihui/bookdown/


## What is LaTeX?  {}

- **Importantly, LaTeX can be included right into RMarkdown documents**
- The following slides have some examples

## Operators and Symbols

```{r, eval=FALSE, echo=TRUE}
$$ \large a^x, \sqrt[n]{x}, \vec{\jmath}, \tilde{\imath}$$
```

$$ \large a^x, \sqrt[n]{x}, \vec{\jmath}, \tilde{\imath}$$

```{r, eval=FALSE, echo=TRUE}
$$ \large \alpha, \beta, \gamma$$
```
$$ \large \alpha, \beta, \gamma$$

## Operators and Symbols

```{r, eval=FALSE, echo=TRUE}
$$ \large\approx, \neq, \nsim $$
```
$$ \large\approx, \neq, \nsim $$

```{r, eval=FALSE, echo=TRUE}
$$\large \partial, \mathbb{R}, \flat$$
```
$$\large \partial, \mathbb{R}, \flat$$

## Equations

Binomial sampling equation

```{r, eval=FALSE, echo=TRUE}
$$\large f(k) = {n \choose k} p^{k} (1-p)^{n-k}$$
```

$$\large f(k) = {n \choose k} p^{k} (1-p)^{n-k}$$

Poisson Sampling Equation

```{r, eval=FALSE, echo=TRUE}
$$\large Pr(Y=r) = \frac{e^{-\mu}\mu^r}{r!}$$
```

$$\large Pr(Y=r) = \frac{e^{-\mu}\mu^r}{r!}$$



## Differential Equations

```{r, eval=FALSE, echo=TRUE}
$$\iint xy^2\,dx\,dy =\frac{1}{6}x^2y^3$$
```
$$\iint xy^2\,dx\,dy =\frac{1}{6}x^2y^3$$

## Matrix formulations

```{r, eval=FALSE, echo=TRUE}
$$	\begin{matrix}
		-2 & 1 & 0 & 0 & \cdots & 0  \\
		1 & -2 & 1 & 0 & \cdots & 0  \\
		0 & 1 & -2 & 1 & \cdots & 0  \\
		0 & 0 & 1 & -2 & \ddots & \vdots \\
		\vdots & \vdots & \vdots & \ddots & \ddots & 1  \\
		0 & 0 & 0 & \cdots & 1 & -2
	\end{matrix} $$
```
	
$$	\begin{matrix}
		-2 & 1 & 0 & 0 & \cdots & 0  \\
		1 & -2 & 1 & 0 & \cdots & 0  \\
		0 & 1 & -2 & 1 & \cdots & 0  \\
		0 & 0 & 1 & -2 & \ddots & \vdots \\
		\vdots & \vdots & \vdots & \ddots & \ddots & 1  \\
		0 & 0 & 0 & \cdots & 1 & -2
	\end{matrix} $$
	
## Including LaTeX and Code into Markdown Files {.flexbox .vcenter}

- Explicit inclusion of code and mathematical equations helps with reproducibility
- Need to designate the 'environment' as being code or math
- Can be included in-line or in 'chunks'

## In-line versus fenced

```{r, eval=FALSE, echo=TRUE}
This equation, $y=\frac{1}{2}$, is included inline
```

This equation, $y=\frac{1}{2}$, is included inline

```{r, eval=FALSE, echo=TRUE}
Whereas this equation, $$y=\frac{1}{2}$$, is put on a separate line
```

Whereas this equation $$y=\frac{1}{2}$$ is put on a separate line

## Markdown is very flexible
- You can import RMarkdown templates into RStudio and open as a new Rmarkdown file
- Better yet there are packages that add functionality
- When you install the package it will show up in the 'From Template' section of the 'new file' startup screen
- There are packages to make
    - books
    - journal articles
    - slide shows
    - interactive exercises
    - many more
- Some of these use 'Shiny'
    - an interactive web based application
    - allows users to input and get output

## Final Thoughts on Markdown, LaTeX and GitHub
- Many forms/flavors of markdown
    - HTML and Rmarkdown are just forms of markdown
    - There is a GitHub flavored markdown
    - Once you learn one, all the others are very easy
- The goal is increased collaboration and reproducibility
    - Allows you to easily work with others by sharing the markdown file
    - Allows **formal** representation of code and math
    - Allows others to run your code directly
    - Allows reports to nontech people
    - All files are easily shared on GitHub
- Once you start using Markdown you won't stop.....